---
title: "Mooseinsitu"
output: html_document
date: "2024-02-11"
---

Load libraries:
```{r libraries}
library(readxl)
library(tidyverse)
library(ggpubr)
library(ggridges)
```

Read in data:
```{r}
act_abund <- read_xlsx("Moose activity and abundance.xlsx", 
                        sheet = 1, col_names = T)
```

set plot colors, etc:
```{r}
palette_treatments <- c("#f0dc82", "#30571f") # gap and closed canopy

```

Plot moose abundance (by droppings) per park, per treatment type:
```{r}
# https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9328498/ (calculating abundance from pellets)

act_abund %>% 
  filter(treatment != "exclosure") %>% 
  ggplot(aes(x=treatment, y = droppings, fill = treatment))+
  geom_boxplot()+
  scale_fill_manual(values=palette_treatments) +
  geom_jitter()+
  theme_classic()+
  facet_wrap(~park)+
  labs(y = "droppings: individual piles")+
  theme(legend.position = "none")
```

Plot moose activity per park, per treatment type:
```{r}
act_abund %>% 
  filter(treatment != "exclosure") %>% 
         # park == "GMNP") %>% 
  ggplot(aes(x=treatment, y = browsing, fill = treatment))+
  # geom_point()+
  geom_boxplot()+
  scale_fill_manual(values=palette_treatments) +
  geom_jitter()+
  theme_classic()+
  ylim(0,80)+
  facet_wrap(~park)+
  labs(y = "browsing: individual plants")+
  theme(legend.position = "none")
```

Merge the above data with the moose density data you gleaned from the Parks: this way, you can associate ambient moose population density with the above in situ observations of activity and abundance. (e.g.: does observed abundance == Parks-counted density? Does activity correlate with Parks-counted density? etc.)
```{r}
# import data generated from Parks Canada maps, with estimates of mean/max moose density at each site sampled in 2023, if not already read in
# tnnp_moose <- read.csv("TNNP_2023sites_moosedensities.csv", header = TRUE)
# gmnp_moose <- read.csv("GMNP_2023sites_moosedensities.csv", header = TRUE)

# append these data to the activity/abundance data: do they match up?
```

