---
title: "moose_density_estimates"
format: html
editor: visual
---

## Moose Density

Estimating moose density across GMNP, TNNP using Rachael's estimates and Parks Canada's survey results.

### libraries:

```{r}
library(tidyverse)
library(readxl)
library(sf)
library(sp)
library(here)
library(ggrepel)
```

# import data:

These data are a combination of Parks Canada data, and from Rachael and/or Shawn.

## Gros Morne NP, Terra Nova NP sampling units: mean and max number of moose observed (density) based on data provided by the parks across multiple sampling years

GMNP: survey units, mean and max moose density over survey years:

```{r GMNP}
# survey units, shape file, GMNP:
su_gmnp <- here("GMNP moose density data", "All GMNP moose survey blocks", "Pre-Stratification_2019.shp")
# make into sf file:
su_gmnp <- st_read(su_gmnp)

# add moose density data: pull maximum total number of moose from the five years' of total numbers in the "su" spreadsheet, calculate density, add to new df:
su_max_gmnp <- su_gmnp %>% 
  pivot_longer(c("X2018_total", 
                 "X2015_total", 
                 "X2017_tot_1",
                 "X2016_total", ), names_to = "totals") %>% 
  group_by(ID) %>% 
  mutate(max = max(value), 
         type = totals[which.max(value)],
         mean = mean(value),
         dens_mean = mean/Area,
         dens_max = max/Area) %>% 
  pivot_wider(everything(), names_from = "totals", values_from = "value")

```

TNNP: survey units, mean and max moose density over survey years:

```{r}
# survey units, KML file, TNNP:
su_tnnp <- ("Moose_Sampling_Units_TNNP.kml")
# make into sf file:
su_tnnp <- st_read(su_tnnp)
# add Sampling Unit (SU) attribute back in:
name <- c(1:82)
su_tnnp$Name <- as.character(name)

# import the density data, which is in a differently-formatted XLSX file in which I just manually calculated mean and max density for all the survey units:
file <- here("rachael_moosedensitydata", "moose survey review 2021_TNNP.xlsx") # paste together parts of the address
tnnp_moose <- read_xlsx(file, sheet = 2)
tnnp_moose$Block <- as.character(tnnp_moose$Block)
# put moose density estimates on the SF file of TNNP:
su_max_tnnp <- left_join(su_tnnp, tnnp_moose, by = join_by("Name" == "Block"))
```

2023 sampling sites, Forbes:

```{r}
gmnp_sites_2023 <- read.csv("2023 Gros Morne.csv", header = T)
gmnp_sites_2023 <- gmnp_sites_2023 %>% 
  # remove some rows that don't have real data
  filter(!row_number() %in% c(7,8,13,14)) 

tnnp_sites_2023 <- read.csv("2023 Terra Nova.csv", header = T)
```

Extract the moose density from the locations where my *2023* sampling sites are: GMNP

```{r}
# project the points from my sites:
# get coordiante reference system (CRS) for gmnp: 
st_crs(su_max_gmnp) # 4269

gmnp_SFpoints_2023 <- data.frame(long = c(gmnp_sites_2023$Longitude), 
                      lat  = c(gmnp_sites_2023$Latitude),
                      site_name = c(gmnp_sites_2023$Name)) %>% 
  st_as_sf(coords = c("long", "lat"), crs = 4269) %>% 
  st_transform(crs = 4269)

# intersect and extract survey unit ID name that encompasses each sampling site of mine: first need to convert the su_max_gmnp geometry to lat/long to match the gmnp_SFpoints_2023 lat/long format geometry
gmnp_su_moosedensity <- st_transform(su_max_gmnp, 4269)

# which survey unit ID (polygon) does each of my sites fall into?
gmnp_sites_density <- st_intersection(gmnp_SFpoints_2023, gmnp_su_moosedensity)

# write.csv(gmnp_sites_density, "GMNP_2023sites_moosedensities.csv")
```

Extract the moose density from the locations where my *2023* sampling sites are: TNNP

```{r}
# project the points from my sites:
# get coordiante reference system (CRS) for gmnp: 
st_crs(su_max_tnnp) # 4326

tnnp_SFpoints_2023 <- data.frame(long = c(tnnp_sites_2023$Longitude), 
                      lat  = c(tnnp_sites_2023$Latitude),
                      site_name = c(tnnp_sites_2023$Name)) %>% 
  st_as_sf(coords = c("long", "lat"), crs = 4326) %>% 
  st_transform(crs = 4326)

tnnp_su_moosedensity <- st_transform(su_max_tnnp, 4326)

# which survey unit ID (polygon) does each of my sites fall into?
sf_use_s2(FALSE) # for some reason, I had to turn off "spherical geometry" to get these two datasets to merge. Not sure why. Going forward with it for now, but making a note that this intersection for TNNP was done with planar assumption of coordinates. (see warning)
tnnp_sites_density <- st_intersection(tnnp_SFpoints_2023, su_max_tnnp)
# although coordinates are longitude/latitude, st_intersection assumes that they are planar
# Warning: attribute variables are assumed to be spatially constant throughout all geometries

# write.csv(tnnp_sites_density, "TNNP_2023sites_moosedensities.csv")
```

# Plotting

## GMNP

```{r gnnp map}

options(ggrepel.max.overlaps = Inf)

gmnp_sites <- su_max_gmnp %>% 
  ggplot() +
  geom_sf(aes(fill = dens_max))+
  geom_sf(data = gmnp_sites_density,
             color = "BLUE", size = 1)+
  # ggrepel::geom_label_repel(
  # data = gmnp_sites_density,
  # aes(label = round(dens_max, 2), geometry = geometry),
  # size = 2,
  # box.padding = 0.1,
  # stat = "sf_coordinates"
  # )+
  scale_fill_continuous(high = "red", low = "white")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90))+
        # legend.position = "None")+
  labs(x = "", y = "", title = "GMNP moose density: max")
gmnp_sites

```

## TNNP

```{r gnnp map}

options(ggrepel.max.overlaps = Inf)

# rename these cols for easier plotting:
su_max_tnnp <- su_max_tnnp %>% 
  rename(avg_moose = "average moose",
         avg_dens = "average density",
         max_moose = "max moose",
         max_dens = "max density",
         area_km2 = "Area (km^2)")

# NOTE: this plot isn't working, I suspect due to the fact that the data aren't aligning in terms of their mismatch in geometry when I had to convert the lat/longs to planar (rather than spherical) when getting the moose densities for my sites in TNNP.
# tnnp_sites <- su_max_tnnp %>% 

# update: the plot wasn't working because the spatial data for TNNP is still 3-D, rather than 2-D (like above when extracting data) so had to two-dimensionalize it for this

# plot 2022 sites only:
tnnp_sites_2022 <- tnnp_sites_density %>% 
  filter(site_name == "TN-excl-ochre" |
           site_name == "TN-Excl-bluehill-E" |
           site_name == "TN-Excl-bluehill-C"
         )

ggplot(st_zm(su_max_tnnp)) +
  geom_sf(aes(fill = max_dens))+
  geom_sf(data = tnnp_sites_density,
             color = "BLUE", size = 1)+
  ggrepel::geom_label_repel(
    data = tnnp_sites_density,
    # data = tnnp_sites_2022,  #making map of 2022 sites alone
    aes(label = round(max.density, 2), geometry = geometry),
    # aes(label = site_name, geometry = geometry), #making map of 2022 sites alone
    size = 2,
    box.padding = 0.1,
    stat = "sf_coordinates")+
  scale_fill_continuous(high = "red", low = "white")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90))+
  labs(x = "", y = "", title = "TNNP moose density: max")

```
