---
title: "vegetationdata_2023"
output: html_document
date: "2023-12-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Vegetation data: collected summer 2023

```{r libraries}
library(tidyverse)
library(readxl)
library(vegan)
```

## upload data:

DBH

```{r data}
dbh <- read_xlsx("Vegetation and soil.xlsx", sheet = 3, col_names = T)
dbh <- dbh[,c(1:8,11,28:29)]
dbh <- dbh %>% 
  rename("grid_loc" = "grid location",
         "DBH_single" = "DBH = CBH/pi",
         "DBH_multi" = "forked only: square root of SS")
dbh <- dbh %>% 
  mutate(net_dbh = coalesce(DBH_single, DBH_multi)) %>% 
  select(c(1:8, net_dbh, 11))

dbh$species <- str_replace_all(dbh$species, "heart-leaved birch", "heartleaf birch")
```

ground vegetation
```{r}
grnd <- read_xlsx("Vegetation and soil.xlsx", sheet = 1, col_names = T)
```

shrubs
```{r}
shrub <- read_xlsx("Vegetation and soil.xlsx", sheet = 2, col_names = T)

shrub <- shrub %>% 
  select(-c(date,patches_heights, `...13`, `...14`, `...15`,
            patches_dimensions, `...18`, `...19`, `...20`))
# need to update these data: homogenize the two different types (= total volume of shrubs per plot)

# clean: we want a "long" dataset that has estimated volume, a note as to whether it's a patch- or an individually-derived estimate, height (again, individual or patch aka average), and area (again, individual or patch derived).
shrub <- shrub %>% 
  mutate(veg_type = if_else((is.na(height)==TRUE), 
                            "patch", "single stem")) %>% #asign vegetation type
  mutate(h_forvol = if_else((is.na(height)==TRUE),
                            avg_h, height)) %>% 
  mutate(area_forvol = )
```

Dead trees
```{r}
dead <- read_xlsx("Vegetation and soil.xlsx", sheet = 4, col_names = T)
```

canopy cover
```{r}
canopy <- read_xlsx("Vegetation and soil.xlsx", sheet = 5, col_names = T)

```

## NFLD palette:
```{r}
palette <- c("#fc9272", "#f0dc82", "#30571f")
```


## visualize data:
ground vegetation:
```{r grnd}
grnd %>% 
  select(-c())
  ggplot(aes(x=treatment, y=TOTAL))+
  geom_boxplot()+
  facet_wrap(~park)+
  theme_classic()+
  theme(legend.position = "none")

# transform grnd data into long format to look at individual groups' cover:
grnd_long <- grnd %>% 
  select(-c(date, notes, site_id, `grid location`, TOTAL)) %>% # don't need these variables
  pivot_longer(!c(site, park, treatment, type), 
               names_to = "functional_group", 
               values_to = "percent_cover")
```
PCA of ground veg: following vegan tutorial here https://rpubs.com/an-bui/vegan-cheat-sheet
```{r}
# isolate just the site x functional groups data from grnd:
grnd_grps <- grnd %>% 
  select(-c("date", "park", "site_id", "treatment", "type", "grid location",
            "TOTAL", "notes"))

grnd_grps$site = as.numeric(as.factor(grnd_grps$site))

pca_grnd <- rda(grnd_grps)
pca_grnd
summary(pca_grnd)
```

Shannon diversity, community-scale perMANOVA (permutational multivariate analysis of variance)
```{r}
# first: set up a metadata frame: with both data (% cover numbers) and metadata associated (treatment, type of gap, etc.)
# count/percent data: the only entries are the values
grnd_percent <- grnd %>% 
  select(-c(date, park, site_id, site, treatment, type, "grid location", TOTAL, notes))
# metadata df:
grnd_env <- grnd %>% 
  select(c(park, site, treatment, type))

# SHANNON DIVERSITY:
shan <- diversity(grnd_percent)
shan_aov <- aov(shan ~ (treatment * park), data = grnd_env)
summary(shan_aov) # significantly different ground veg communities in the park, and the treatments in each park

# perMANOVA null: centroids of group (aka similiarity) are equivalent across groups
perman <- adonis2(grnd_percent ~ (treatment*park), data = grnd_env)
perman # ground cover differs by park, treatment, and by the interaction

# PCA: get ordination of each functional group in "variance" space (how much influence does it have on community compostion as it pertains to the variables we're interested in)
grnd_pca <- rda(grnd_percent)

# PLOT:
pca_scores <- scores(grnd_pca, display = "sites") %>% 
  as.data.frame() %>% 
  # rownames_to_column("site") %>% 
  cbind(grnd_env)

pca_vect <- scores(grnd_pca, display = "species") %>% 
  as.data.frame()

pca_scores %>% 
  ggplot() +
  geom_point(aes(x = PC1, y = PC2, color = treatment)) +
  scale_color_manual(values = palette) +
  geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
  geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +
  geom_segment(data = pca_vect, aes(x = 0, y = 0, xend = PC1, yend = PC2), 
               arrow = arrow(length = unit(0.2, "cm"))) +
  geom_text(data = pca_vect, aes(x = PC1, y = PC2, label = rownames(pca_vect))) +
  labs(x = "PC1",
       y = "PC2") +
  theme_classic()
```
NMDS plot:
```{r}
grnd_nmds <- metaMDS(grnd_percent)
grnd_nmds # low stress on the model, this is good: 0.1737944 (aka the ordination did a pretty good job representing community structure)
stressplot(grnd_nmds)

plot_df <- scores(grnd_nmds, display = "sites") %>% 
  as.data.frame() %>% 
  # rownames_to_column("site") %>% 
  cbind(grnd_env)

ggplot(plot_df, aes(x = NMDS1, y = NMDS2, color = treatment, shape = treatment)) +
  geom_point(size = 3, alpha = 0.8) +
  scale_color_manual(values = palette) +
  stat_ellipse(linetype = 2, size = 1) +
  theme_classic() +
  labs(title = "NMDS plot of ground cover communities")
```

# canopy:
```{r}
# Tidy the data:
canopy <- canopy %>% 
  rename(paper_birch = `paper birch (% of closed)`,
         balsam_fir = `balsam fir (% of closed)`,
         american_ash = `American ash (% of closed)`,
         speckled_alder = `Speckled alder (% of closed)`,
         black_spruce = `Black spruce (% of closed)`,
         red_maple = `Red Maple (% of closed)`,
         larch = `Larch (% of closed)`,
         tamarack = `tamarack (% of closed)`)
canopy <- canopy %>% 
  mutate_at(c("paper_birch", "balsam_fir", "american_ash",
              "speckled_alder", "black_spruce", "red_maple",
              "larch", "tamarack"), as.numeric) %>% 
  mutate_at(c("paper_birch", "balsam_fir", "american_ash",
              "speckled_alder", "black_spruce", "red_maple",
              "larch", "tamarack"), ~na_if(., 0))

canopy %>% 
  ggplot(aes(x=treatment, y=closed))+
  geom_boxplot(aes(fill = treatment))+
  scale_fill_manual(values = palette)+
  ylab("percent of canopy that is closed")+xlab("")+
  theme_classic()+
  theme(legend.position = "none")

# convert to long-form
canopy_long <- canopy %>% 
  select(-c(date, site_id)) %>% # don't need these variables
  pivot_longer(!c(site, park, treatment, type, `grid location`, open, closed), 
               names_to = "tree_spp", 
               values_to = "percent_cover") %>% 
  mutate(percent_cover = as.numeric(percent_cover))
```

canopy cover: exclosure vs. mature
```{r}
canopy_long %>% 
  filter(treatment != "gap") %>% 
  ggplot(aes(x=treatment, y=closed))+
  geom_boxplot()+
  theme_classic()
  
```
Canopy cover: exclosure vs. mature, NMDS/PCA
```{r}
# first: set up a metadata frame: with both data (% cover numbers) and metadata associated (treatment, type of gap, etc.)
# count/percent data: the only entries are the values
canopy_percent <- canopy %>% 
  select(-c(date, park, site_id, site, treatment, type,
            "grid location", open, closed))
canopy_percent <- canopy_percent %>% 
  mutate(across(where(is.numeric), tidyr::replace_na, 0))
  
# metadata df:
canopy_env <- canopy %>% 
  select(c(park, site, treatment, type, "grid location"))

# SHANNON DIVERSITY:
shan <- diversity(canopy_percent)
shan_aov <- aov(shan ~ (treatment * park), data = canopy_env)
summary(shan_aov) # significantly different across treatments but not parks

# perMANOVA null: centroids of group (aka similiarity) are equivalent across groups
perman <- adonis2(canopy_percent ~ (treatment*park), data = canopy_env)
perman # ground cover differs by park, treatment, and by the interaction

# PCA: get ordination of each functional group in "variance" space (how much influence does it have on community compostion as it pertains to the variables we're interested in)
canopy_pca <- rda(canopy_percent)

# PLOT:
pca_scores <- scores(canopy_pca, display = "sites") %>% 
  as.data.frame() %>% 
  # rownames_to_column("site") %>% 
  cbind(canopy_env)

pca_vect <- scores(canopy_pca, display = "species") %>% 
  as.data.frame()

pca_scores %>% 
  ggplot() +
  geom_point(aes(x = PC1, y = PC2, color = treatment)) +
  scale_color_manual(values = palette) +
  geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
  geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +
  geom_segment(data = pca_vect, aes(x = 0, y = 0, xend = PC1, yend = PC2), 
               arrow = arrow(length = unit(0.2, "cm"))) +
  geom_text(data = pca_vect, aes(x = PC1, y = PC2, label = rownames(pca_vect))) +
  labs(x = "PC1",
       y = "PC2") +
  theme_classic()
```
NMDS plot:
```{r}
canopy_nmds <- metaMDS(canopy_percent)
canopy_nmds # low stress on the model, this is good: 0.1737944 (aka the ordination did a pretty good job representing community structure)
stressplot(canopy_nmds)

plot_df <- scores(grnd_nmds, display = "sites") %>% 
  as.data.frame() %>% 
  # rownames_to_column("site") %>% 
  cbind(grnd_env)

ggplot(plot_df, aes(x = NMDS1, y = NMDS2, color = treatment, shape = treatment)) +
  geom_point(size = 3, alpha = 0.8) +
  scale_color_manual(values = palette) +
  stat_ellipse(linetype = 2, size = 1) +
  theme_classic() +
  labs(title = "NMDS plot of ground cover communities")
```




```{r hist}
dbh %>% 
  # filter(treatment == "gap") %>% 
  ggplot(aes(x=net_dbh, group = park, fill = park))+
  # ggplot(aes(x=net_dbh, group = species, fill = species))+
  # geom_density(adjust=1.5, alpha = 0.75)+
  geom_density(alpha = 0.75)+
  scale_fill_manual(values = c("darkolivegreen", "goldenrod"))+
  labs(y="", x="diameter at breast height (cm)")+
  theme_classic()
```
