---
title: "Moose density data"
output: html_notebook
---

This script pulls moose population density data from existing Parks Canada records to go with the transect datasets I collected in summer 2022 (for which I have GPS data points).  The idea is to associate a moose population density with each site if possible. (Ideal scenario would be to have data on moose presence / absence inside and out of the canopy gaps, but these data are too coarse for that.)

Librarys:
```{r libraries}
library(tidyverse)
library(readxl)
library(sf)
library(here)
library(ggrepel)
```

Read in moose data (from parks):
```{r readdat}
here()
su <- here("GMNP moose density data", "All GMNP moose survey blocks", "Pre-Stratification_2019.shp")
fixed <- here("GMNP moose density data", "fixed moose monitoring plots", "monitoring_plots24.shp")
fname <- here("GMNP moose density data", 
              "Fixed moose survey block results 2007-2021", 
              "MonitoringSurvey_DensityResults2007-2021.shp")
moose <- st_read(fname)
su <- st_read(su)
fixed <- st_read(fixed)

View(moose)

moose_geom <- st_geometry(moose)

moose_plot <- sf::st_transform(moose)

plot(moose[,"Density_16"], reset = F)
```

What I want to do is determine in which "SU" (survey unit) each of my sampling sites falls into. I will assign each one an SU, then pull the moose density from that SU into the df of my own sampled data.
1) plot the SUs
2) plot my sampling sites on top of the SU's
3) assign an SU to each sampling site
4) determine moose density for each SU and thus each sampling site

Add data from fluxdat4: lat/long. these data are in A DIFFERENT crs (coordinate reference system) than the moose dataset from Parks Canada.  I took those lat/long data by hand, in WGS84 crs (reference code = 4326). The moose data is in NAD83; these two CRS's are similar, but not quite the same (which is why when i plot them without specifying the CRS as the same, it looks like I sampled right off the coast of Rocky Harbour).
```{r}
# make compressed df with lats/longs of sites, not each collar:
sites <- fluxdat5 %>% 
  group_by(site_name) %>% 
  summarise(lon = mean(latitude),
            lat = mean(longitude)) # reminder that lat/lon are mixed up in my orig dataset

gmnp_sites <- sites %>% 
  filter(lat > 49)  # filter out the TNNP sites

moose_plot %>% 
  ggplot() +
  geom_sf(aes(fill=Density_16)) +
  annotate("point", x = c(gmnp_sites$lon), 
           y = c(gmnp_sites$lat), colour = "red", size = 0.5) +
  coord_sf(default_crs = sf::st_crs(4326))+
  theme_classic()
  # theme(legend.position = "none")

```
Merge the two datasets to pull the survey unit ID and match it to the coordinates of each site of mine (e.g. which SU was each site located in?).
```{r merge_1}

moose_plot %>% 
  ggplot() +
  geom_sf(aes(fill = Density_16)) +
  geom_sf(data = su, fill = NA, color = "lightgrey")+
  geom_text_repel(data = gmnp_sites, aes(x = lon, y = lat, label = site_name),
                  size = 4, col = "darkred",
                  nudge_x = c(-.2,-.2,-.3,0, -0.2, .7),
                  nudge_y = c(.1,.2,.4,-.1,-.2,-.1))+
  annotate("point", x = c(gmnp_sites$lon),
           y = c(gmnp_sites$lat), colour = "red", size = 0.5) +
  coord_sf(default_crs = sf::st_crs(4326))+
  theme_classic()+
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(angle = 20, vjust = -0.01))+
  theme(axis.title = element_blank())
```

List the total number of moose in each SU over time:
```{r SU_obs}

su %>% 
  ggplot() +
  geom_sf(aes(fill = X2018_total))+
  theme_classic()+
  theme(legend.position = "none")

# pull maximum total number of moose from the five years' of total numbers in the "su" spreadsheet, calculate density, add to new df:
su_max <- su %>% 
  pivot_longer(c("X2018_total", 
                 "X2015_total", 
                 "X2017_tot_1",
                 "X2016_total", ), names_to = "totals") %>% 
  group_by(ID) %>% 
  mutate(max = max(value), 
         type = totals[which.max(value)],
         mean = mean(value),
         dens_mean = mean/Area,
         dens_max = max/Area) %>% 
  pivot_wider(everything(), names_from = "totals", values_from = "value")
```

```{r map}
# project the points from my sites:
crs_use <- "+proj=laea +lat_0=30 +lon_0=-95"
d_points <- data.frame(long = c(gmnp_sites$lon), 
                      lat  = c(gmnp_sites$lat),
                      site_name = c(gmnp_sites$site_name)) %>% 
  st_as_sf(coords = c("long", "lat"), crs = 4326) %>% 
  st_transform(crs = crs_use)

# plot density with my sites overlaid:
su_max_labs <- su_max %>% filter(dens_mean > 0) %>% mutate(dens_r = round(dens_mean, digits = 1))
su_max %>% 
  ggplot() +
  geom_sf(aes(fill = dens_mean))+
  # annotate("point", x = c(gmnp_sites$lon),
           # y = c(gmnp_sites$lat))+
  # geom_sf(data = d_points, 
             # color = "BLUE", size = 1)+
  geom_sf_label(data = su_max_labs, 
                aes(label = dens_r), size = 2, 
                label.padding = unit(0.1, "lines"))+
  scale_fill_continuous(high = "red", low = "white", na.value="lightyellow")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 20, vjust = -0.01))+
  labs(x = "", y = "")
```

Read in xlsx data from drive folder:
```{r xlsx}
moose_dat_all <- read_excel("2019_GMNP moose survey data.xlsx", sheet = "park-wide")

# join these data to the one with the SU locations:
moosedat_merge <- left_join(su_max, moose_dat_all, by = "ID") 

# plot density with my sites overlaid:
moosedat_merge %>% 
  ggplot() +
  geom_sf(aes(fill = (Total_moose/Area)))+
  # geom_sf(data = d_points, fill = "darkred",
             # color = "red", size = 1)+
  scale_fill_gradient(high = "darkred", low = "pink", na.value = "white")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 20, vjust = -0.01))

# intersect and extract SU ID name:
d_points_trans <- st_transform(d_points, 4326) # apply transformation to pnts sf
moosedat_trans <- st_transform(moosedat_merge, 4326)      # apply transformation to polygons sf

d_points$SU_ID <- apply(st_intersects(moosedat_trans, d_points_trans, sparse = FALSE), 2, 
               function(col) { 
                  moosedat_merge[which(col), ]$ID
               })
d_points$density <- apply(st_intersects(moosedat_trans, d_points_trans, sparse = FALSE), 2,
                          function(col){
                            moosedat_merge[which(col), ]$dens
                          })
View(d_points)

```

Plot using the data from above:
```{r}
d_points <- st_transform(d_points, 4326)      # apply transformation to polygons sf

moosedat_merge %>% 
  ggplot() +
  geom_sf(aes(fill = dens))+
  geom_sf(data = d_points,
             color = "red", size = 1)+
  scale_fill_continuous(high = "darkred", low = "white", na.value = "white")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 20, vjust = -0.01))
```
One more datasheet:
```{r}
moose_2009_2021 <- read_csv(here("GMNP moose density data", "moose_monitoring_fixed blocks_alldata_2009to2021.csv"))

# join these data to the one with the SU locations:
moosedat_merge2 <- left_join(su_max, moose_2009_2021, by = "ID")

# plot density with my sites overlaid:
moosedat_merge2 %>% 
  filter(dens < 3.1) %>% 
  ggplot() +
  geom_sf(aes(fill = Density))+
  geom_sf(data = d_points, fill = "darkred",
             color = "red", size = 1)+
  scale_fill_gradient(high = "darkred", low = "lightblue", na.value = "white")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 20, vjust = -0.01))
```

##################################
18 March 2023: got new density data from Shawn.
##################################
##################################
26 May 2023: Rachael estimated moose density data
##################################

Import Rachael's data:
```{r}
rachael_moose <- read_xlsx(here("rachael_moosedensitydata", "All_sites(New).xlsx"))
r_gmnp <- rachael_moose %>% 
  filter(Park == "GM")
r_gmnp <- data.frame(long = c(r_gmnp$Lat), 
                      lat  = c(r_gmnp$Long),
                      site_name = c(r_gmnp$Site)) %>% 
  st_as_sf(coords = c("lat", "long"), crs = 4326) %>% 
  st_transform(crs = crs_use)

# plot density with my sites overlaid:
su_max %>% 
  ggplot() +
  geom_sf(aes(fill = (dens_mean)))+
  geom_sf(data = r_gmnp, color = "red", size = 1)+ # add Rachael's sites
  geom_sf(data = d_points, color = "darkred", size = 1)+ # add my sites
  geom_sf_label(data = su_max_labs, 
                aes(label = dens_r), size = 2, 
                label.padding = unit(0.1, "lines"))+
  scale_fill_gradient(high = "orange", low = "white", na.value = "white")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 20, vjust = -0.01))+
  guides(fill=guide_legend(title="density (per sq. km)"))

# intersect and extract SU ID name:
d_points_trans <- st_transform(d_points, 4326) # apply transformation to pnts sf
moosedat_trans <- st_transform(moosedat_merge, 4326)      # apply transformation to polygons sf

d_points$SU_ID <- apply(st_intersects(moosedat_trans, d_points_trans, sparse = FALSE), 2, 
               function(col) { 
                  moosedat_merge[which(col), ]$ID
               })
d_points$density <- apply(st_intersects(moosedat_trans, d_points_trans, sparse = FALSE), 2,
                          function(col){
                            moosedat_merge[which(col), ]$dens
                          })
View(d_points)
```


